---
title: "Classical Growth Analysis (Hunt et al., 2002)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Classical Growth Analysis (Hunt et al., 2002)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(phytogrowR)
library(dplyr)
```

## 1. Introduction

Classical plant growth analysis summarizes interval-average indices between
successive harvests. In contrast, functional analysis fits curves and derives
instantaneous rates. This vignette follows the practical recommendations from
Hunt et al. (2002).

## 2. Required data

For the classical module, the main variables are:

- total biomass
- leaf biomass
- root biomass
- shoot biomass (or leaf + stem [+ reproductive biomass])
- leaf area
- harvest time

## 3. Example dataset

```{r}
dplyr::glimpse(hunt_classical_example)
```

## 4. Classical calculation with `classical_growth_interval()`

```{r}
fit_classic <- classical_growth_interval(
  data = hunt_classical_example,
  time = time,
  total_biomass = total_biomass_g,
  leaf_biomass = leaf_biomass_g,
  root_biomass = root_biomass_g,
  stem_biomass = stem_biomass_g,
  reproductive_biomass = reproductive_biomass_g,
  leaf_area = leaf_area_cm2,
  group_by = treatment,
  ci_method = "both",
  n_boot = 199,
  seed = 123
)

print(fit_classic)
```

```{r}
summary(fit_classic)
```

```{r}
head(tidy(fit_classic))
```

## 5. Interpretation of indices

- `RGR`: interval mean relative growth.
- `ULR/NAR`: interval mean assimilation per unit leaf area (operationally equivalent here).
- `SLA`: area per unit leaf biomass.
- `LWF/LWR`: leaf biomass fraction (synonyms in this module).
- `LAR`: area per unit total biomass, with decomposition `LAR = SLA * LWF`.
- `root_shoot_ratio`: root allocation relative to shoot.
- `root_shoot_allometric_coefficient`: slope (`beta`) from `log(root) ~ log(shoot)`.

## 6. Treatment comparison

```{r}
fit_treatment <- tidy(fit_classic) %>%
  filter(index %in% c("rgr", "ulr", "sla", "lar")) %>%
  arrange(index, treatment)

fit_treatment
```

## 7. Graphics

```{r}
plot(fit_classic, metric = "rgr")
```

```{r}
plot(fit_classic, metric = "ulr")
```

```{r}
plot_allometry(
  data = hunt_classical_example,
  shoot_biomass = "shoot_biomass_g",
  root_biomass = "root_biomass_g",
  treatment = "treatment"
)
```

## 8. Limitations

- Classical indices are interval means, not instantaneous rates.
- Exactly two harvests are required per interval.
- Log-based indices require strictly positive biomass and leaf area.
- When many harvests are available, functional analysis may be preferable.

## 9. References

Hunt, R., Causton, D. R., Shipley, B., & Askew, A. P. (2002). A modern tool for classical plant growth analysis. *Annals of Botany*, 90(4), 485-488. doi:10.1093/aob/mcf214
