---
title: "Functional Growth Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functional Growth Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(phytogrowR)
library(dplyr)
```

## Why Functional Growth Analysis?

Classical metrics are interval means. Functional methods estimate smooth trajectories and derive instantaneous rates from derivatives.

## Fit Curves

```{r}
dat <- prepare_growth_data(growth_wide_example)
fit_w <- fit_growth_curve(dat, response = "total_biomass_g", group_cols = c("treatment"), method = "gam")
fit_la <- fit_growth_curve(dat, response = "leaf_area_cm2", group_cols = c("treatment"), method = "gam")

summary(fit_w)
plot_growth_curve(fit_w)
```

## Instantaneous Rates

```{r}
inst <- calc_instant_rates(fit_w, leaf_area_fit = fit_la, epsilon = 1e-6)
head(inst)
plot_growth_rates(inst, type = "instantaneous", metrics = c("agr_inst", "rgr_inst", "algr_inst", "rlgr_inst", "nar_inst"))
```

## Compare Treatments

```{r}
cmp_rgr <- compare_growth(dat, metric = "rgr", group_var = "treatment", epsilon = 1e-6)
cmp_rgr

boot_rgr <- bootstrap_growth(dat, metric = "rgr", group_var = "treatment", times = 100, seed = 2026, epsilon = 1e-6)
boot_rgr
```

## Formal Test of Coincidence of Curves

`compare_growth()` contrasts one index at a time. When the question is whether
the treatments follow *the same growth trajectory*, the comparison should be
made between whole curves, and that is a nested-model problem:

* **H0 (reduced model)** — one single curve describes every treatment, so the
  parameter vector is shared: \(p_0 = p\) parameters.
* **H1 (full model)** — every treatment keeps its own parameter vector:
  \(p_1 = k\,p\) parameters.

`compare_growth_curves()` fits both models and compares them with the F test
for nested models and with the likelihood ratio test:

\[
F = \frac{(RSS_0 - RSS_1)/(p_1 - p_0)}{RSS_1/(n - p_1)},
\qquad
G = n \log\!\left(\frac{RSS_0}{RSS_1}\right) \sim \chi^2_{p_1 - p_0}.
\]

Using `degree = 2` on the log scale is the classical log-quadratic functional
growth analysis, where the first derivative of the fitted curve is the relative
growth rate:

```{r}
ct <- compare_growth_curves(
  dat,
  response = "total_biomass_g",
  group_var = "treatment",
  model = "poly",
  degree = 2,
  log_response = TRUE,
  epsilon = 1e-6,
  pairwise = TRUE
)

ct
```

The `plot()` method makes the quantity being tested visible: the coloured lines
are the treatment-specific curves and the dashed black line is the single common
curve imposed by H0.

```{r}
plot(ct)
```

Besides the global `coincidence` hypothesis, the polynomial route splits the
difference into `level` (equal intercepts given a common shape) and `shape`
(parallel curves). The non-linear families instead report one hypothesis per
parameter, which tells you *why* the curves differ:

```{r}
nl <- compare_growth_curves(
  dat,
  response = "total_biomass_g",
  group_var = "treatment",
  model = "logistic"
)

tidy(nl)
```

A significant `Asym` row points to different final sizes, `xmid`/`b` to
different timing, and `scal`/`c`/`k`/`r` to different rates. Because a single
residual variance is assumed for all treatments, check the residuals before
trusting a borderline p-value; heteroscedastic biomass data usually calls for
`log_response = TRUE`.

## Agronomic Interpretation

- `AGR` and `CGR` describe absolute biomass accumulation speed.
- `RGR` and `NAR` help interpret physiological efficiency through time.
- Partition metrics highlight allocation shifts under stress versus nutrient supply.
- `ALGR`, `RLGR`, and `LAI` strengthen interpretation of canopy expansion dynamics.

## Classical References

- Hunt, R. (1990). *Basic Growth Analysis*.
- Benincasa, M. M. P. (2003). *Plant Growth Analysis: Basic Concepts*.

